---
title: "Descriptive Statistics"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

Download the script [here](https://raw.githubusercontent.com/LuMesserschmidt/stats_TUM/master/Files/Course%20Scripts/descriptives.R)


***

# Preliminary

* We want to analyze our data. In the first place this means we are going to conduct descriptive analyses: How is the data structured, what do we know about the variables? 

* To do so, we need to download the data, do we?

***


This is how the beginning of every of your scripts should look like: 1. Clean environment and 2. load packages  
```{r, message=FALSE}
rm(list = ls())
# install.packages("tidyverse")
library(tidyverse)
```

***

# Loading the dataset

To make life easier, we load a dataset that is already attached to R: the mtcar dataset.

```{r, eval=FALSE}
data<- mtcars
```

## Explore the data structure

For a first overview over our data, we can use the base functions (e.g. those which are without any package installations part of R) `names()`,`str()`, `head()`, `summary()`, `table()`, `quantile()`, and `View()`.

<span style="color:red">Task</span> <u>**Go through every function's output and try to understand what it means!**</u>
```{r, eval=FALSE}
names(data)
str(data)
head(data)
summary(data)
table(data$mpg) # you could use any other variable here.
quantile(data$mpg, na.rm = TRUE)
View(data)
```

***
# Data Distribution
In the beginning of each research project, we start with most important data indicators: We want to know the `mean()`, the `median()`, or about the `range()` of a variable. 

```{r, eval=FALSE}
mean(data$cyl, na.rm=T) 
median(data$cyl, na.rm=T) 
range(data$cyl, na.rm=T)
```

***
# Frequency Table

## One-dimensional

Or we want to check about the frequency: We use the `table()` function.

To give an example:

```{r, eval=FALSE}
table(data$cyl)
```

## n-dimensional table

Sometimes, it is interesting whether there are differences among categories. For example, we want to know the two-diemsnional table for the number of gears and cylinders.

```{r, eval=FALSE}
table(data$cyl,data$gear)
```

# Histogram

Moving away from relative or absolute probabilities, we can analyse the distribution. Histograms are most common for this:

```{r, eval=FALSE}
hist(data$gear, main = paste("Gear Histogram"),xlab="Cylinder in numbers", ylab="Frequence")
```


# Boxplot

Yes, R is pretty straightforward. For example, we can easily calculate the boxplot for the variables of any data: 

```{r, eval=FALSE}
boxplot(data$mpg, na.rm=T,main=paste("Boxplot mpg"))
```



